Solutions to Integer Programming from Quantum Annealing

### 2020-09-10 paper/main.tex
Added more context to first result section and simplified qubo notation

In this paper, we propose a new algorithm that maps ILP problems to QUBOs.
This is achieved by the introduction of slack variables $s$ which turn the inequalities eq.~\eqref{eq:ilp-constraints} to equalities
\begin{align}
  & \sum_i A_{a i}x_i + s_a + b_a = 0, \\
  & s_a \geq 0,                \\
  & x_i \in \mathbb{Z} \geq 0.
\end{align}

@@

%----------------------------------------------------------------------------------------
\subsection{ILP QUBO}
\label{sec:results:qa1}
%----------------------------------------------------------------------------------------
{\color{red} This section and the methods section reads very disjoint. You also don't define everything here and then Reference the Methods for more information.}

The integer quadratic optimization problem can be mapped to a minimization of the quadratic objective function
\begin{align}
 \chi^2                         & = \left[\Psi^x_{\mu} T^x_{\mu i}d_{ij} T^x_{j \nu}\Psi^x_\nu + c_i T^x_{i\mu} \Psi^x_\mu \right.
 \nonumber                                                                                                                             \\ \label{eq:qubo-min}
                                & \qquad\left.+ p (A_{a b} T^x_{b \mu} \Psi^x_{\mu} + T^s_{a \alpha} \Psi^s_\alpha + b_a)^2 \right],   \\
 \vec x_0                       & =  T^x \vec \Psi^x_0 \, ,
 \\
 (\vec \Psi^x_0, \vec \Psi^s_0) & = \mathrm{arg}\min\limits_{\vec{\Psi}}(\chi^2) \,,                                                 &
\end{align}
where $p$ is an external parameter representing the strength of the penalty when violating constraints and $T^{x,z}$ is a matrix mapping qubits to integers (see also section \ref{sec:methods:bit-basis}).
The objective function can be represented as a QUBO Hamiltonian
\begin{align}
 E =    &
 \begin{pmatrix}
  \Psi^x & \Psi^s
 \end{pmatrix}
 \begin{pmatrix}
  Q_{xx} & Q_{xs} \\
  Q_{sx} & Q_{ss}
 \end{pmatrix}
 \begin{pmatrix}
  \Psi^x \\ \Psi^s
 \end{pmatrix} + pb^2, \\
 \equiv & \Psi^T Q \Psi + C,
 \label{eq:matrix_form}
\end{align}
where
\begin{align}
 Q_{xx} = & {T^{x}}^T \left[ d + p A^T A + p \mathrm{Diag}_{x} \left(A^T b + b^T A\right) \right] T^x \nonumber \\
          & + \mathrm{Diag}_{\psi^x}(c) T^x,                                                                    \\
 Q_{xs} = & Q_{sx}^T = p {T^{x}}^T A^T T^s,                                                                     \\
 Q_{ss} = & p\left[ {T^{s}}^T T^s + \mathrm{Diag}_{\psi^s}\left( {T^{s}}^T b + b^T T^s\right) \right].
\end{align}
The function $\mathrm{Diag}_{z}(v)$ embeds a vector $v$ into a diagonal matrix in subspace $z$, and absorbs the linear contributions of the QUBO into the diagonal elements of the quadratic representation.

A solution to the ILP is found by obtaining the qubit vector $\psi$ which minimizes $E$.
The integer solution to the original ILP is computed by $\vec x = T^x \vec \psi^x$ and the original energy is computed by shifting the annealer extracted energy by $pb^2$.
The slack components of this vector $\psi^s$ can be utilized to check if the contraints are indeed fulfilled.

@@

The solution to the minimum independent set can be expressed as an integer optimization problem given by,
\begin{align}
 f(x) =                    & \min\left(\sum_{\nu \in V} \psi^x_{\nu}\right),                    \\
 \textrm{subject to} \quad & \psi^x_{\nu} + \sum_{\mu \in \mathit{N}(\nu)} \psi^x_{\mu} \geq 1, \\
                           & \psi^x_{\nu} \in \{0, 1\}
\end{align}
where the dimension of the dependent variable $x$ is the number of vertices $\overline{\overline{V}}$.
The problem minimizes the number of vertices in $D$ with a binary variable $\psi^x$ encoded by a single qubit, subject to the constraint that at least one vertex in $\mathit{N}(\nu)$ is in $D$.
For each vertex in $V$ we introduce slack variables $s_{\nu} = T^s_{\nu \alpha} \psi^s_{\alpha}$ to encode the inequality constraint such that
\begin{align}
 f(x) =                    & \min(\sum_{\nu\in V} \psi^x_{\nu}),                                                 \\
 \textrm{subject to} \quad & \psi^x_{\nu} + J_{\nu \mu} \psi^x_{\mu}- T^s_{\nu \alpha} \psi^s_{\alpha}  - 1 = 0, \\
                           & \psi_{\nu} \in \{0, 1\},                                                            \\
                           & 0 \leq s_{\nu} \leq |\mathrm{N}(\nu)|                                               \\
                           & s_{\nu} \in \mathbb{Z},
\end{align}
where the nearest-neighbor sum is expressed by $J$ (symmetric and zero diagonal), the adjacency matrix for $G$.
The algorithm uses $N_q = \overline{\overline{V}} + \sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits---$\overline{\overline{V}}$ qubits to encode the vertices and another $\sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits to embed the slack variables.
Therefore, the embedding at worse scales with $\overline{\overline{V}} \log_2 |\overline{\overline{V}}|$ qubits for fully connected graphs.

The target QUBO in the notation of eq.~\eqref{eq:matrix_form} reads
 {\small
  \begin{align}
   Q_{xx} & = \mathbbm{1} + p\left[J^T J + J^T + J - 2 \mathrm{Diag}_{\psi^x}(|J|) - \mathbbm{1} \right] \,, \\
   Q_{sx} & = - p(\mathbbm{1}+J^T)T^s\,,                                                                     \\
   Q_{ss} & = p\left[{T^s}^T T^s + 2\mathrm{Diag}_{\psi^s}(|T^s|)\right]\,,                                  \\
   C      & =  p \overline{\overline{V}}\,,
  \end{align}}
where $ |M| \equiv \sum_{\nu} M_{\nu \mu}$.

@@

%----------------------------------------------------------------------------------------
\subsubsection{{\color{green} Integer quadratic optimization}}
\label{sec:methods:ilp:notes}
%----------------------------------------------------------------------------------------

@@

While the coefficients of the inequality constraints are not required to be integer valued, the inequalities can be trivially rescaled such that $s_i \in \mathbbm{N}$ given fixed precision coefficients $A_{ij}$ and $b_i$.

>>>>

In this paper, we propose a new algorithm that maps ILP problems to QUBOs.
This is achieved by the introduction of slack variables $s$ which turn the inequalities eq.~\eqref{eq:ilp-constraints} to equalities
\begin{align}
  \label{eq:ilp:slack}
  & \sum_i A_{a i}x_i + s_a + b_a = 0, \\
  & x_i, s_a \in \mathbb{Z} \geq 0.
\end{align}
While the coefficients of the inequality constraints are not required to be integer valued, the inequalities can be trivially rescaled such that $s_i \in \mathbbm{Z}$ given fixed precision coefficients $A_{ij}$ and $b_i$.
Furthermore, this formalism can be generalized to constrained quadratic optimizations \ref{sec:methods:ilp:quadratic}.

@@

%----------------------------------------------------------------------------------------
\subsection{ILP QUBO}
\label{sec:results:qa1}
%----------------------------------------------------------------------------------------
We map the integer variables $\vec z = \vec x, \vec s$ appearing in eq.~\eqref{eq:ilp:slack} to qubits under the following transformation~\cite{Chang:2018uoc}
\begin{align}
 z_i = & \sum_{r=0}^{R_i-1} 2^r \psi_{ri}
 \label{eq:int_to_bin}
\end{align}
where $\psi_{ri} \in \{0, 1\}$.
The number of qubits used to represent the $i$-th integer variable is allowed to vary with $R_i$ (see also section \ref{sec:methods:bit-basis}).
This transformation can be expressed as a rectangular matrix $T_z$ such that
\begin{equation}
    \begin{pmatrix}
        \vec x \\ \vec s
    \end{pmatrix}
    =
    \begin{pmatrix}
        T_x & 0 \\ 0 & T_s
    \end{pmatrix}
    \begin{pmatrix}
        \vec \Psi_x \\ \vec \Psi_s
    \end{pmatrix}
    =
    T
    \vec \Psi
\end{equation}

The integer linear (and quadratic equivalently, see section \ref{sec:methods:ilp:quadratic}) optimization problem can be mapped to a minimization of the quadratic objective function
\begin{align}
 \chi^2(\vec \Psi)
 &=
 \vec c^T T_x \vec \Psi_x + p \left\| A T_s \vec \Psi_x + T_s \vec \Psi_s + \vec b \right\|^2
\end{align}
where $p$ is an external parameter representing the strength of the penalty when violating constraints.
This parameter needs to be sufficiently large, e.g., $p \geq \vec c \cdot \vec x$, to ensure the constraints are indeed fulfilled\footnote{Depending on the problem, $p$ can be smaller as well.
For example, in the case of the MDS, $p\geq 1$ suffices.}.
The objective function $\chi^2$ can be represented as a QUBO Hamiltonian
\begin{align}
 \chi^2(\vec \Psi) =    &
 \vec \Psi^T
 \begin{pmatrix}
  Q_{xx} & Q_{xs} \\
  Q_{sx} & Q_{ss}
 \end{pmatrix}
 \vec \Psi + p\left \| \vec b \right\|^2, \\
 \equiv &  \vec \Psi^T Q  \vec \Psi + C,
 \label{eq:matrix_form}
\end{align}
where
{\small
\begin{align}
 \label{eq:qubo:components}
 Q_{xx} = & p {T_{x}}^T \left[ A^T A + \mathrm{diag} \left(A^T \vec b + \vec b^T A\right) \right] T_x + \mathrm{diag}(\vec c) T_x,                                                                    \\
 Q_{xs} = & Q_{sx}^T = p {T_{x}}^T A^T T_s,                                                                     \\
 Q_{ss} = & p\left[ {T_{s}}^T T_s + \mathrm{diag}\left( {T_{s}}^T \vec b + \vec b^T T_s\right) \right].
\end{align}}
The function $\mathrm{diag}(\vec v)$ transforms a vector $\vec v$ into a diagonal matrix, and absorbs the linear contributions of the QUBO into the diagonal elements of the quadratic representation.

A solution to the ILP is found by identifying the qubit vector $\vec \Psi^{(0)}$ which minimizes $ \chi^2$.
The integer solution to the original ILP is computed by $\vec x^{(0)} = T_x \vec \Psi_x^{(0)}$ and the original solution to the problem is computed by shifting the annealer extracted energy $E \equiv  \chi^2(\vec \Psi^{(0)}) $ by $p \left \| \vec b \right\|^2$.
The slack components of this vector $\vec \Psi_s$ can be utilized to check if the contraints are indeed fulfilled.

@@

The solution to the minimum independent set can be expressed as an integer optimization problem given by,
\begin{align}
 f(\vec x) = & \min\left(\sum_{v \in V} x_v\right),                    \\
\end{align}
subject to
\begin{align}
 & x_v + \sum_{j \in \mathcal{N}(v)} x_j \geq 1, \\
 & x_v \in \{0, 1\}
\end{align}
where $\mathcal{N}(v)}$ represents the set of all direct neighbors of vertex $v$ and the dimension of the dependent variable $x$ is the number of vertices $n_V \equiv \overline{\overline{V}}$.
The problem minimizes the number of vertices in $D$ with a binary variable $\vec x = \vec \Psi_x$ encoded by single qubit, subject to the constraint that at least one vertex in $\mathcal{N}(\nu)$ is in $D$.
For each vertex in $V$ we introduce slack variables
\begin{equation}
    \vec s \in \left\{ \mathbb{Z}^{n_V} \, \middle| \, 0 \leq s_{v} \leq |\mathcal{N}(v)| \quad \forall v\in V \right\} \, ,
\end{equation}
which is related to the qubit vector $\vec \Psi_s$ by $\vec s = T_s \vec \Psi_s$, to encode the inequality constraint such that
\begin{align}
 f(\vec \Psi_x)
 =
 & \min\left(\vec 1 \cdot \vec \Psi_x \right) \, ,
\end{align}
subject to
\begin{align}
 &
 \vec \Psi_x + J \vec \Psi_x - T_s \Psi_s  - \vec 1 = 0\,,
 \\
 &
 (\vec \Psi_s)_\nu \in \{ 0, 1\}
\end{align}
where the nearest-neighbor sum is expressed by $J$ (zero diagonal and symmetric for non-directional graphs), the adjacency matrix for $G$.
The algorithm uses
\begin{equation}
    N_q = \overline{\overline{V}} + \sum_{\nu \in V} \log_2 \mathit(N(\nu))
\end{equation}
qubits to encode the vertices and to embed the slack variables.
Therefore, the embedding at worse scales with $n_V \log_2 n_V$ qubits for fully connected graphs.

The target QUBO in the notation of eq.~\eqref{eq:matrix_form} reads
 {\small
  \begin{align}
   Q_{xx} & = \mathbbm{1} + p\left[J^T J + J^T + J - 2 \mathrm{diag}(|J|) - \mathbbm{1} \right] \,, \\
   Q_{sx} & = - p(\mathbbm{1}+J^T)T_s\,,                                                                     \\
   Q_{ss} & = p\left[{T_s}^T T_s + 2\mathrm{diag}(|T_s|)\right]\,,                                  \\
   C      & =  p \overline{\overline{V}}\,,
  \end{align}}
where $ |M| \equiv \sum_{\nu} M_{\nu \mu}$.

@@

%----------------------------------------------------------------------------------------
\subsubsection{Integer quadratic optimization}
\label{sec:methods:ilp:quadratic}
%----------------------------------------------------------------------------------------

@@

In this case the $Q_{xx}$ component of the Qubo, eq.~\eqref{eq:qubo:components} obtains a new term
\begin{equation}
    Q_{xx} \to Q_{xx} + T_x^T d T_x \, .
\end{equation}
